\chapter*{Abstract}
\label{cha:abtract}
\addcontentsline{toc}{chapter}{Abstract}

The proliferation of embedded devices across industries and daily life, from consumer
electronics to healthcare and industrial automation, has amplified the urgency
of securing these systems against an escalating array of cyber threats. This thesis
presents \textit{STEERED: Secure Trusted Execution Environment for RISC-V-based Embedded
Devices}, a novel security framework designed to protect embedded systems leveraging
the open \textit{RISC-V} Instruction Set Architecture. Unlike proprietary ISAs,
\textit{RISC-V} enables the development of customized processors and solutions tailored
to specific use cases. \textit{STEERED} exploits this flexibility to address control-flow
tampering attacks, such as Return-Oriented Programming and Jump-Oriented
Programming, which threaten the integrity of execution paths in microcontroller-based
devices.

\textit{STEERED} incorporates lightweight Control Flow Integrity enforcement
mechanisms, including a Shadow Stack and a Control Flow Graph, to ensure that
software execution adheres strictly to predefined paths. This prevents unauthorized
deviations caused by potential attackers, ensuring the security of both the
system and its data. The solution employs code instrumentation techniques that automatically
integrate CFI into user applications, simplifying deployment and reducing
development time. Moreover, the use of \textit{RISC-V}'s Physical Memory Protection
safeguards critical data structures, enhancing the robustness of the framework.

The design philosophy behind \textit{STEERED} emphasizes efficiency and scalability,
enabling it to operate effectively on resource-constrained embedded devices. A key
feature is its ability to provide strong security guarantees with minimal performance
overhead, making it suitable for IoT devices that often have limited
computational and power resources. A proof-of-concept implementation demonstrates
the effectiveness of \textit{STEERED} in defending against control-flow
hijacking attacks. Detailed performance analysis reveals that the framework achieves
a balance between security and operational efficiency, with an evaluation of
memory overhead and execution time supporting its feasibility for practical
applications.

Furthermore, this thesis explores the integration of \textit{STEERED} with Real-Time
Operating Systems such as \textit{FreeRTOS} and \textit{Zephyr RTOS}, showcasing
its adaptability to complex software environments. This integration broadens the
applicability of \textit{STEERED}, enabling the secure execution of real-time
and multitasking applications in IoT and embedded domains.

In conclusion, \textit{STEERED} addresses the pressing challenge of securing
embedded \textit{RISC-V} devices by combining innovative security mechanisms,
automated processes, and adherence to \textit{RISC-V}'s open and customizable
principles. This work contributes significantly to the field of embedded system security
by offering a scalable, efficient, and customizable framework that meets the needs
of modern IoT environments while anticipating future challenges in the cybersecurity
landscape.