\chapter{Conclusions}
\label{cha:conclusions}

This thesis discussed the design, development, and analysis of a Control Flow
Integrity enforcer for \textit{RISC-V}-based embedded devices which poses itself
as a significant contribution to enhancing security in resource-constrained environments.
The project successfully addresses key challenges related to embedded device
security, including mitigating control flow hijacking attacks such as Return-Oriented
Programming and Jump-Oriented Programming. With the integration of advanced
mechanisms like a shadow stack and Control Flow Graph, the project ensures robust
enforcement of Control Flow Integrity, a critical step in preventing
unauthorized execution flows or arbitrary code execution.

In chapters \ref{cha:background} and \ref{cha:threatmodel}, we provided a comprehensive
description of control flow hijacking attacks describing hwo they work and why they
pose such a threat to modern-day security of embedded systems. Moreover, we described
the currently available security measures to mitigate this class of cyber threats.
Lastly, we discussed the project's threat model, highlighting the scope and
limitations of the provided security features.

In chapter \ref{cha:project} we have seen the design and implementation
processes highlighting their key features. We have also seen how the project adheres
to the principles of lightweight and efficient security, making it suitable for
devices with limited computational resources. The system leverages the flexibility
and extensibility of the \textit{RISC-V} Instruction Set Architecture to create
a tailored solution that balances security and performance. The proposed infrastructure
achieves this balance through meticulous code instrumentation, optimized
Physical Memory Protection configurations, and efficient trap management mechanisms.
Additionally, the modular design of the project ensures that it remains adaptable
to various use cases and allows for further customization. Lastly, with the Proof
of Concept presented in section \ref{sec:project_poc} we demonstrated the project's
ability to effectively detect and mitigate attacks that aim at disrupting the code
execution flow.

Chapter \ref{cha:ta} was focused on providing an in-depth view of the testing methodologies
used during the project's security assessment. Additionally, we proved how it can
effectively protect the code from control flow hijacking attacks. Also, we discussed
the weak points of the project, describing its limitations in protecting from
specific attacks and discussing potential enhancements to provide a higher
degree of security.

A comprehensive performance analysis presented in chapter \ref{cha:pa}
illustrates the practical viability of incorporating the project into real-world
applications while maintaining a high level of system efficiency. The analysis delves
into the average overhead associated with the implementation, highlighting the
importance of this metric in evaluating its effectiveness. We explored various optimization
techniques specifically designed to mitigate elevated overheads that may arise in
certain unusual scenarios. By addressing these potential challenges, we aim to ensure
that the proposed solution can be seamlessly integrated into existing systems,
ultimately enhancing overall performance while minimizing any adverse effects.

The purpose of chapter \ref{cha:rtos} was to provide an implementation idea to
integrate the system with widely used Real-Time Operating Systems such as
\textit{FreeRTOS} and \textit{Zephyr RTOS}. Such integration could enhance the project's
versatility and readiness for broader adoption, allowing for the execution of
more complex tasks. We focused on this implementation because many real-world applications
are highly dependent on the features provided by RTOSes to carry out their tasks.

Lastly, the thesis also identifies several areas for future exploration, including
the incorporation of additional security mechanisms, further code optimization, and
exhaustive testing under diverse operational conditions in chapter \ref{cha:future}.
These potential enhancements aim to refine the system and broaden its
applicability, extending its impact on the security of embedded systems in the rapidly
expanding Internet of Things landscape.

In conclusion, the project discussed in this thesis marks a significant
milestone in the realm of embedded system security. This innovative approach introduces
practical solutions specifically designed to protect \textit{RISC-V}
microcontrollers from a variety of emerging cyber threats that have been
increasingly targeting these systems. By addressing not only the immediate security
challenges faced by these microcontrollers but also laying the groundwork for future
advancements, it represents a forward-thinking initiative in cybersecurity.

A key achievement is the focus on safeguarding untrusted code from control-flow
hijacking attacks, a common and dangerous threat in the embedded systems landscape.
The developed infrastructure has been meticulously designed to ensure that
security measures do not detrimentally affect the performance of resource-constrained
systems. This balance between security and efficiency is critical, as many
embedded devices operate under stringent resource limitations.

Furthermore, this work highlights the increasing necessity of secure and efficient
design principles within embedded device infrastructures. As embedded systems
continue to integrate deeper into various aspects of modern technology
ecosystems, ranging from consumer electronics to critical infrastructure, the
importance of robust security mechanisms cannot be overstated. The findings and advancements
brought forth by this project not only enhance the immediate security posture of
\textit{RISC-V} microcontrollers but also serve as a foundational framework for
ongoing innovation in the field, ultimately contributing to a safer and more secure
technological landscape.
